---
title: "Latin Americans Interest in Politics vs. Satisfaction with Government"
author: "Karen Perez Torres SIS-750"
date: "2024-03-27"
output:
    beamer_presentation:
    theme: "Malmoe"
    colortheme: "Rochester"
fontsize: 10pt
classoption: "aspectratio=169"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = 'center',
  include = TRUE
)
```

```{r}
library(tidyverse)
library(ggplot2)
library(readxl)
library(haven)
library(patchwork)
```


```{r}
values<-read.csv("WVScsv.csv")

#Change over time of Domocratic Opinions in Latin America
values2 <- filter(values, B_COUNTRY == 68 | B_COUNTRY == 32 | B_COUNTRY == 76 | B_COUNTRY == 152 | B_COUNTRY == 170 |B_COUNTRY == 218 | B_COUNTRY == 484 | B_COUNTRY == 558 | B_COUNTRY == 604 | B_COUNTRY == 320 | B_COUNTRY == 862 | B_COUNTRY == 858) #Latin America

#Democracy
#Q250: How Important is Democracy?
#Q251: How democratically is this country being governed today?
#Q252: Satisfaction with the political system performance?

values3 = values2 |>
  group_by(A_YEAR) |>
  summarize(avg_250 = mean(Q250, na.rm = TRUE),
            avg_251 = mean(Q251, na.rm = TRUE),
            avg_252 = mean(Q252, na.rm = TRUE)
  )


#Nicaragua Political Importance vs. Satisfaction with Governing System in 2020?
awareness <- filter(values, B_COUNTRY == 558 & Q252 >0 & Q4 > 0)

#Q4: Important in life: Politics (For each of the following aspects, indicate how important it is in your life. 
#Would you say it is very important, rather important, not very important or not important at all?)
##Q252: Satisfaction with the political system performance?
awareness2 = awareness |>
  group_by(Q4) |>
  summarize(avg_252 = mean(Q252, na.rm = TRUE)
  )
#Visualization would be a bar chart of each response (1 = Very, 2=Kind of) with its avg Q252 response

#FINDING: In Nicaragua, people are pulled to the extremes. Either they are very engaged AND Satisfied
# OR they are not engaged AND dissatisfied. WHY? Hypothesis: Totalitarianism
  
#Linear Regression of Political Importance to Satisfaction for Latin America
awarenessla <- filter(values2, Q252 >0 & Q4 > 0)
latinawareness <- lm(Q252 ~ Q4, data = awarenessla)

#Theme of Graph
Theme <- theme_minimal() +
  theme(
    text = element_text(),
    plot.title = element_text(face = 'bold', hjust = 1),
    plot.subtitle = element_text(hjust = 1),
    plot.background = element_rect(fill = 'cornflowerblue', color = 'NA'),
    legend.position = 'center',
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x = element_line(color = 'black'),
    axis.text.y = element_text(color = 'black')
  )
Theme2 <- theme_minimal() +
  theme(
    text = element_text(),
    plot.title = element_text(face = 'bold', hjust = 1),
    plot.subtitle = element_text(hjust = 1),
    plot.background = element_rect(fill = 'white', color = 'NA'),
    legend.position = 'center',
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x = element_line(color = 'black'),
    axis.text.y = element_text(color = 'black')
  )
#Bar Graph
P1 = awareness2 |>
  ggplot(aes(
    x = as.factor(Q4), 
    y = avg_252,
    fill = as.factor (Q4),
  )) +
  geom_col() +
  labs(
    title = 'Fig.1',
    x = "Are politics important in your life?",
    y = "Avg. Satisfaciton in Political System",
    subtitle = "Satisfaction with the political performance in Nicaragua",
  ) +
  scale_fill_manual(
    values = c("1" = "white", "2" = "white", "3" = "white", "4" = "white")
  ) +
  Theme

#Nicaragua 

P2 = awareness |>
  ggplot(aes(
    x = Q4,
    y = Q252,
  )) +
  geom_jitter(shape = 11, color = "cornflowerblue", alpha = .2 ) +
  geom_smooth(method = lm, se = TRUE, color = 'black') +
  scale_x_continuous(name = "Satisfied with Political System") +
  scale_y_continuous(
    name = "Importance of Politics") +
    labs(title = "Fig.2") +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.minor = element_blank(),
    legend.margin = margin(0, 0, 0, 20),
    legend.position = "top",
    legend.box = "horizontal") +
  Theme2

#Latin America

P3 = awarenessla |>
  ggplot(aes(
    x = Q4,
    y = Q252,
  )) + 
  geom_jitter(shape = 10, color = "red", alpha = .2 ) +
  geom_smooth(method = lm, se = TRUE, color = 'black') +
  scale_x_continuous(name = "Satisfied with Political System") +
  scale_y_continuous(
    name = "Importance of Politics") +
  labs(title = "Fig.3") +
  theme_minimal(base_size = 8) +
  theme(
    panel.grid.minor = element_blank(),
    legend.margin = margin(0, 0, 0, 20),
    legend.position = "top",
    legend.box = "horizontal",
    legend.box.spacing = grid::unit(0, "pt"),
    legend.justification = c(1, 0),
    legend.spacing.x = unit(2, "pt"),
    legend.spacing.y = unit(2, "pt"),
    legend.background = element_rect(fill = "white", color = NA),
    legend.key.width = unit(10, "pt")
  )
```

##  Data: The World Values Survey Association (WVS)
- Comprised data for 64 countries/territories. 
- The 7th wave of the WVS took place worldwide in 2017-2022.
- 353 Questions
- Political Interest and Political Participation

Source: https://www.worldvaluessurvey.org/WVSDocumentationWV7.jsp

---

## 1 - Nicaragua 
- Fig. 1 Key:
(1=Very Important 4=Not at all Important)

  (1=Not Satisfied At All 10=Completely Satisfied)

```{r fig.width = 5, fig.height = 2.675}
plot(P1)
``` 
---

## 2 - Relationship between the importance of politics and satisfaction with the political system in Nicaragua

```{r fig.width = 5, fig.height = 2.675}
plot(P2)
```

---

## 3 - Relationship between the importance of politics and satisfaction with the political system in Latin America

```{r fig.width = 4, fig.height = 2.675}
plot(P3)
```

---

## FINDINGS:

In Latin America as a whole, people are pretty neutral on satisfaction, and lower engagement doesn't cause huge declines in sanctification.

Future Questions: Do we see similar trends with totalitarians around the world? 
Could we group by totalitarian/non-totalitarian and run comparisons?
